---
title: "CDRscope: Interpretable TCR/BCR Repertoire Analysis"
author: "CDRscope Authors"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{CDRscope Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4.2, fig.align = "center",
                      message = FALSE, warning = FALSE)
```

## Overview

`CDRscope` integrates the six research directions of immune-repertoire
biology into a single interpretable workflow:

1. **Input** — `ReadRepertoire()` / `fetch_vdjdb()` / `fetch_toy_data()`
2. **QC** — `QCRepertoire()` / `NormalizeRepertoire()`
3. **Features** — `ComputeFeatures()` (six concept modules)
4. **Embedding** — `ConceptBottleneckEmbed()` (concept bottleneck)
5. **Classification** — `DiseaseClassify()` (linear + SHAP)
6. **Markers** — `FindMarkers()` (statistical + sequence)

Every step preserves biological semantics, so the disease-separation result is
traceable back to interpretable concept axes.

```{r load}
library(CDRscope)
```

## 1. Build a toy repertoire

```{r toy}
obj <- fetch_toy_data(n_samples = 40, n_clones_per_sample = 300, seed = 1)
obj
```

## 2. QC and normalise

```{r qc}
obj <- QCRepertoire(obj, min_len = 6, max_len = 30)
obj <- NormalizeRepertoire(obj, method = "freq")
obj@misc$qc
```

## 3. Compute six-concept features

```{r features}
obj <- ComputeFeatures(obj)
dim(obj@features)
obj@feature_modules
```

## 4. Concept-bottleneck embedding

```{r embed}
obj <- ConceptBottleneckEmbed(obj, method = "pca", ndim = 10)
DimPlot(obj)
```

## 5. Disease classification with attribution

```{r classify}
obj <- DiseaseClassify(obj, method = "logistic", use_shap = TRUE)
obj@classification$train_accuracy
SHAPPlot(obj)
```

## 6. Biomarker discovery

```{r markers}
obj <- FindMarkers(obj, level = "both")
head(obj@markers$statistical_sig, 8)
head(obj@markers$sequence_sig, 8)
```

## 7. Visualisation gallery

```{r vis1}
FeaturePlot(obj, "convergence")
```

```{r vis2}
VlnPlot(obj, "shannon")
```

```{r vis3}
FeatureHeatmap(obj)
```

```{r vis4, fig.height=4}
MotifLogo(obj, group = "infection", top_n = 15)
```

## Interpretability

Each plot maps back to a named concept axis: the concept-bottleneck layer
forces every disease decision through the six interpretable modules, and SHAP
attributes the decision to specific axes. Statistical markers localise *which
mechanism is perturbed*; sequence markers give *concrete CDR3 motifs* as
candidate biomarkers.
